% vim: spl=en_us
\section{Spectra of operators in Banach algebras}
As we have introduced the set of invertible operators of a unital associative algebra, we may study the notion of spectra and resolvent sets. In Banach algebras, the spectrum of an operator is much simpler than in general algebras, as we will show. The results here derived will be particularized to bounded operators in Banach or Hilbert spaces at a later point; for now we focus solely on Banach algebras.
\begin{definition}{Resolvent and spectrum}{resolvent_spectrum}
    Let \(\algebra{A}\) be a unital associative algebra and let \(a \in \algebra{A}\). The set
    \begin{equation*}
        \rho(a) = \setc*{\lambda \in \mathbb{C}}{\lambda \unity - a \in \invertible{\algebra{A}}}
    \end{equation*}
    is the \emph{resolvent set of \(a\)}.  The set
    \begin{equation*}
        \sigma(a) = \setc*{\lambda \in \mathbb{C}}{\lambda \unity - a \notin \invertible{\algebra{A}}}
    \end{equation*}
    is the \emph{spectrum \(\sigma(a)\) of \(a\)}.
\end{definition}
\begin{remark}
    It is clear the spectrum is the complement of the resolvent set, that is, \(\sigma(a) = \mathbb{C} \setminus \rho(a)\).
\end{remark}

\cref{prop:prop411} can be stated in terms of the spectra and resolvent sets of \(uv\) and \(vu\).
\begin{corollary}
    Let \(\algebra{A}\) be a unital associative algebra. Let \(u, v \in \algebra{A}\), then \(\sigma(uv) \setminus \set{0} = \sigma(vu) \setminus\set{0}\) and \(\rho(uv)\setminus\set{0} = \rho(vu)\setminus\set{0}\).
\end{corollary}
\begin{proof}
    Let \(\lambda \in \rho(vu)\setminus\set{0}\), then \(\lambda \unity - vu \in \invertible{\algebra{A}}\). By \cref{prop:invertible_scalar}, we have \(\unity - \lambda^{-1} vu \in \invertible{\algebra{A}}\), hence \( \unity - \lambda^{-1} uv \in \invertible{\algebra{A}}\) by \cref{prop:prop411}, that is, \(\rho(vu) \setminus \set{0} \subset \rho(uv)\setminus \set{0}\). The converse is shown by replacing \(u\) and \(v\).

    Let \(\lambda \in \sigma(uv) \setminus \set{0},\) then \(\lambda \unity - uv \notin \invertible{\algebra{A}}\). \cref{prop:invertible_scalar} shows that \(\unity - \lambda^{-1} uv\) must not be invertible. \cref{prop:prop411} lets us conclude \(\unity - \lambda^{-1} vu \notin \invertible{\algebra{A}}\), hence \(\lambda \in \sigma(vu) \setminus\set{0}\). The converse is shown \emph{mutatis mutandis}.
\end{proof}

That is, the spectra of \(uv\) and \(vu\) may only be different at zero, which we illustrate with the shift operators in \(\ell_2\).
\begin{example}{Spectra of the shift operators}{spectra_shift}
    Consider the shift operators \(S, S^* \in \bounded(\ell_2)\) defined in \cref{exam:shift_example}. Then \(\sigma(S\circ S^*) = \set{0,1}\) and \(\sigma(S^* \circ S) = \set{1}\).
\end{example}
\begin{proof}
    It is clear \(\rho(S^*\circ S) = \mathbb{C} \setminus \set{1}\), since \(\lambda\unity - S^* \circ S = (\lambda - 1)\unity\), which is invertible for all \(\lambda \neq 1\). To see \(\set{0,1}\subset\sigma(S \circ S^*)  \), we consider the sequences \(u,v \in \ell_2\) defined by
    \begin{equation*}
        u(n) = \begin{cases}
            1,&\text{if }n = 1\\
            0,&\text{otherwise}.
        \end{cases}
        \quad\text{and}\quad
        v(n) = \begin{cases}
            0,&\text{if }n = 1\\
            2^{-n},&\text{otherwise}
        \end{cases}
    \end{equation*}
    Then, \(S \circ S^*u = 0\) and \((\unity - S \circ S^*)v = 0\), hence \(S \circ S^*\) and \((\unity - S\circ S^*)\) are not injective. The previous corollary guarantees the spectrum of \(S\circ S^*\) is \(\set{0,1}\).
\end{proof}

An immediate consequence of the previous corollary is the similarity invariance of the spectrum.
\begin{proposition}{Similarity invariance}{similarity_invariance}
    Let \(\algebra{A}\) be a unital associative algebra. If \(u \in \invertible{\algebra{A}}\), then \(\sigma(uvu^{-1}) = \sigma(v)\) for all \(v \in \algebra{A}\).
\end{proposition}
\begin{proof}
    It is clear that \(\sigma(uvu^{-1})\setminus\set{0} = \sigma(v)\setminus\set{0}\). Since
    \begin{equation*}
        0 \in \rho(v) \iff v \in \invertible{\algebra{A}} \iff uvu^{-1} \in \invertible{\algebra{A}} \iff 0 \in \rho(uvu^{-1}),
    \end{equation*}
    we have \(\sigma(uvu^{-1}) = \sigma(v)\).
\end{proof}
\begin{corollary}
    If \(u,v \in \invertible{\algebra{A}}\), then \(\sigma(uv) = \sigma(vu)\).
\end{corollary}
\begin{proof}
    We have \(\sigma(u(vu)u^{-1}) = \sigma(vu)\) by the previous proposition.
\end{proof}

The spectra of the adjoint and of the inverse operators to an invertible operator can be easily related to its spectrum.
\begin{proposition}{Spectrum of the inverse operator}{spectrum_inverse}
    Let \(\algebra{A}\) be a unital associative algebra. If \(u \in \invertible{\algebra{A}}\), then \(\sigma(u^{-1}) = \setc{\lambda \in \mathbb{C}}{\lambda^{-1} \in \sigma(u)}\).
\end{proposition}
\begin{proof}
    Since \(u \in \invertible{\algebra{A}}\), \(0 \notin \sigma(u)\). For \(\lambda \in \mathbb{C} \setminus \set{0}\), we have
    \begin{equation*}
        \lambda \unity - u^{-1} \notin \invertible{\algebra{A}} \iff -\lambda^{-1} u \left(\lambda \unity - u^{-1}\right) \notin \invertible{\algebra{A}} \iff \lambda^{-1}\unity - u \notin \invertible{\algebra{A}},
    \end{equation*}
    hence \(\lambda \in \sigma(u^{-1})\) if and only if \(\lambda^{-1} \in \sigma(u)\).
\end{proof}
\begin{proposition}{Spectrum of the adjoint of an operator}{spectrum_adjoint}
    Let \(\algebra{A}\) be a unital *-algebra. If \(u \in \invertible{\algebra{A}}\), then \(\sigma(u^*) = \setc{\lambda \in \mathbb{C}}{\conj{\lambda} \in \sigma(u)}\).
\end{proposition}
\begin{proof}
    Since \(\invertible{\algebra{A}}\) is self-adjoint, we have
    \begin{equation*}
        \lambda \unity - u^* \notin \invertible{\algebra{A}} \iff (\lambda \unity - u^*)^* \notin \invertible{\algebra{A}} \iff \conj{\lambda}\unity - u \notin \invertible{\algebra{A}},
    \end{equation*}
    hence \(\lambda \in \sigma(u^*)\) if and only if \(\conj{\lambda} \in \sigma(u)\).
\end{proof}
For an invertible operator \(u \in \invertible{\algebra{A}}\) in a unital *-algebra \(\algebra{A}\), we'll denote
\begin{equation*}
    \sigma(u)^{-1} = \setc{\lambda \in \mathbb{C}}{\lambda^{-1} \in \sigma(u)} \quad\text{and}\quad \conj{\sigma(u)} = \setc{\lambda \in \mathbb{C}}{\conj{\lambda} \in \sigma(u)}
\end{equation*}
for the spectra of \(u^{-1}\) and \(u^*\), respectively.

\subsection{Topological properties of the spectrum}
We may associate an operator with each complex number in the resolvent set of an operator.
\begin{definition}{Resolvent operator}{resolvent_operator}
    Let \(\algebra{A}\) be a unital associative algebra. Let \(a \in \algebra{A}\) and \(\lambda \in \rho(a)\), the operator
    \begin{equation*}
        R_\lambda(a) = (\lambda \unity - a)^{-1}
    \end{equation*}
    is the \emph{resolvent of \(a\) at \(\lambda\)}.
\end{definition}

By studying the resolvent operator, we may derive many conclusions about the spectra of operators. We begin with simple properties of the resolvent.
\begin{proposition}{Resolvent and operator commute}{resolvent_commute}
    Let \(\algebra{A}\) be a unital associative algebra. If \(R_{\lambda}(a) \in \algebra{A}\) is the resolvent of \(a \in \algebra{A}\) at \(\lambda \in \rho(a)\), then \(R_{\lambda}(a)a = aR_{\lambda}(a)\).
\end{proposition}
\begin{proof}
    Since \(R_{\lambda}(a) \in \invertible{\algebra{A}}\), we have
    \begin{align*}
        R_{\lambda}(a)^{-1} R_{\lambda}(a) = R_{\lambda}(a) R_{\lambda}(a)^{-1} = \unity
        &\implies  \lambda R_{\lambda(a)} - aR_{\lambda}(a) = \lambda R_{\lambda(a)} - R_{\lambda}(a) a\\
        &\implies a R_{\lambda}(a) = R_{\lambda}(a)a
    \end{align*}
    as desired.
\end{proof}

\begin{proposition}{First resolvent identity}{first_resolvent_identity}
    Let \(\algebra{A}\) be a unital associative algebra and let \(a \in \algebra{A}\). Then
    \begin{equation*}
        R_{\lambda}(a) - R_{\mu}(a) = (\mu - \lambda)R_{\lambda}(a)R_{\mu}(a)
    \end{equation*}
    for all \(\lambda, \mu \in \rho(a)\).
\end{proposition}
\begin{proof}
    We have
    \begin{equation*}
        R_{\lambda}(a) = R_{\lambda}(a)(\mu \unity - a)R_{\mu}(a) = R_{\lambda}(a) \left[(\mu - \lambda)\unity + (\lambda \unity - u)\right]R_{\mu}(a) = (\mu - \lambda)R_{\lambda}(a) R_{\mu}(a) - R_{\mu}(a)
    \end{equation*}
    for all \(\lambda, \mu \in \rho(a)\).
\end{proof}
\begin{corollary}
    Let \(\algebra{A}\) be a unital associative algebra and let \(a \in \algebra{A}\). Then, \(R_{\lambda}(a)R_{\mu}(a) = R_{\mu}(a)R_{\lambda}(a)\) for all \(\lambda, \mu \in \rho(a)\).
\end{corollary}
\begin{proof}
    The first resolvent identity yields
    \begin{equation*}
        (\mu - \lambda)R_{\lambda}(a)R_{\mu}(a) = R_{\lambda}(a) - R_{\mu}(a) = -(\lambda - \mu)R_{\mu}(a)R_{\lambda}(a),
    \end{equation*}
    then
    \begin{equation*}
        (\mu - \lambda)\left[R_{\lambda}(a) R_{\mu}(a) - R_{\mu}(a) R_{\lambda}(a)\right] = 0
    \end{equation*}
    for all \(\mu,\lambda \in \rho(a)\) and the result follows.
\end{proof}
\begin{proposition}{Second resolvent identity}{second_resolvent_identity}
    Let \(\algebra{A}\) be a unital associative algebra and let \(u, v \in \algebra{A}\). Then,
    \begin{equation*}
        R_{\lambda}(u) - R_{\lambda}(v) = R_{\lambda}(u) (u - v) R_{\mu}(v)
    \end{equation*}
    for all \(\lambda \in \rho(u) \cap \rho(v)\).
\end{proposition}
\begin{proof}
    We have
    \begin{equation*}
        R_{\lambda}(u) (u-v) R_{\lambda}(v) = R_{\lambda}(u) \left[(\lambda \unity - v) - (\lambda \unity - u)\right] R_{\lambda}(v) = R_{\lambda}(u) - R_{\lambda}(v)
    \end{equation*}
    for all \(\lambda \in \rho(u) \cap \rho(v)\).
\end{proof}
\begin{corollary}
    Let \(\algebra{A}\) be a unital associative algebra and let \(u, v \in \algebra{A}\). Then,
    \begin{equation*}
        R_{\lambda}(u) - R_{\lambda}(v) = R_{\lambda}(v) (u - v) R_{\mu}(u)
    \end{equation*}
    for all \(\lambda \in \rho(u) \cap \rho(v)\).
\end{corollary}
\begin{proof}
    Exchanging \(u\) and \(v\) in the second identity yields the result.
\end{proof}
\begin{proposition}{Third resolvent identity}{third_resolvent_identity}
    Let \(\algebra{A}\) be a unital associative algebra and let \(u, v \in \algebra{A}\). Then,
    \begin{equation*}
        R_{\lambda}(uv) = \lambda^{-1}\left[\unity + uR_{\lambda}(vu)v\right]
    \end{equation*}
    for all \(\lambda \in \rho(vu)\setminus\set{0}\).
\end{proposition}
\begin{proof}
    Let \(\lambda \in \rho(vu)\setminus\set{0},\) then \(R_{\lambda}(vu) = \lambda^{-1}(\unity - \lambda^{-1}vu)^{-1}\). We repeat the proof of \cref{prop:prop411}:
    \begin{align*}
        \left[\lambda^{-1}\unity + \lambda^{-1}uR_{\lambda}(vu)v\right](\lambda\unity - uv)
        &= \unity + uR_{\lambda}(vu)v - \lambda^{-1}uv - \lambda^{-1}uR_{\lambda}(vu)vuv\\
        &= \unity - \lambda^{-1}uv + uR_{\lambda}(vu)(\unity - \lambda^{-1}vu)v\\
        &= \unity - \lambda^{-1}uv + \lambda^{-1}uv\\
        &= \unity,
    \end{align*}
    and
    \begin{align*}
        (\lambda\unity - uv)\left[\lambda^{-1}\unity + \lambda^{-1}uR_{\lambda}(vu)v\right]
        &= \unity - \lambda^{-1}uv + uR_{\lambda}(vu)v - \lambda^{-1}uvuR_{\lambda}(vu)v\\
        &= \unity - \lambda^{-1}uv + u(\unity - \lambda^{-1}vu)R_{\lambda}(vu)v\\
        &= \unity - \lambda^{-1}uv + \lambda^{-1}uv\\
        &= \unity,
    \end{align*}
    proving our claim.
\end{proof}

\begin{lemma}{Resolvent set is open}{resolvent_set_open}
    Let \(\algebra{A}\) be a unital Banach algebra and let \(a \in \algebra{A}\). If \(\mu \in \rho(a)\), then the open ball \(B_{\norm{R_\mu(a)}^{-1}}(\mu)\) is contained in \(\rho(a)\) and
    \begin{equation*}
        R_{\lambda}(a) = R_\mu(a)\left\{\unity + \sum_{n = 1}^{\infty} \left[(\mu - \lambda)R_\mu(a)\right]^n\right\} = \left\{\unity + \sum_{n = 1}^{\infty}\left[(\mu - \lambda)R_\mu(a)\right]^n\right\}R_{\mu}(a)
    \end{equation*}
    for all \(\lambda \in B_{\norm{R_\mu(a)}^{-1}}(\mu)\).
\end{lemma}
\begin{proof}
    Notice
    \begin{equation*}
        (\kappa \unity - a)R_\mu(a) = \left[(\kappa - \mu)\unity + (\mu \unity - a)\right]R_{\mu}(a) = \unity + (\kappa - \mu)R_{\mu}(a),
    \end{equation*}
    for all \(\kappa \in \mathbb{C}\).

    Let \(\lambda \in B_{\norm{R_\mu(a)}^{-1}}(\mu)\), then \(\norm{(\mu - \lambda)R_{\mu}(a)} < 1\), hence \(\unity - (\mu - \lambda)R_{\mu}(a) \in \invertible{\algebra{A}}\) and
    \begin{equation*}
        \unity + \sum_{n = 1}^{\infty} \left[(\mu - \lambda)R_\mu(a)\right]^n = \left[\unity - (\mu - \lambda)R_{\mu}(a)\right]^{-1} = \left[(\lambda \unity - a)R_\mu(a)\right]^{-1}
    \end{equation*}
    by \cref{thm:neumann_series_algebra}. Since \(\invertible{\algebra{A}}\) is closed under the product and \(R_\mu(a) \in \invertible{\algebra{A}}\), we have \(\lambda \in \rho(a)\) and
    \begin{equation*}
        \unity + \sum_{n = 1}^{\infty} \left[(\mu - \lambda)R_\mu(a)\right]^n = (\mu\unity - a)R_{\lambda}(a),
    \end{equation*}
    concluding our proof.
\end{proof}
\begin{corollary}
    Let \(\algebra{A}\) be a unital Banach algebra. The spectrum of an operator is a closed subset of \(\mathbb{C}\).
\end{corollary}
\begin{proof}
    Let \(a \in \algebra{A}.\) If \(\rho(a)\) is empty, then \(\sigma(a) = \mathbb{C}\) is closed, so we may assume \(\rho(a)\) is non-empty. Since every point of \(\rho(a)\) is an interior point by \cref{lem:resolvent_set_open}, we have \(\mathbb{C} \setminus \rho(a)\) open.
\end{proof}

\begin{lemma}{Linear functionals define holomorphic maps}{holomorphic}
    Let \(\algebra{A}\) be a unital Banach algebra and \(a \in \algebra{A}\). Let \(\ell \in \algebra{A}^\dag\) be a continuous linear functional, then the map
    \begin{align*}
        f_{\ell} : \rho(a) &\to \mathbb{C}\\
                   \lambda &\mapsto \ell(R_\lambda(a))
    \end{align*}
    is holomorphic in each connected component of \(\rho(a)\).
\end{lemma}
\begin{proof}
    Let \(\mu \in \rho(a)\) and consider \(\lambda \in B_{\norm{R_\mu(a)}^{-1}}(\mu)\setminus \set{\mu},\) then
    \begin{equation}
        f_\ell(\lambda) = \ell(R_{\lambda}(a)) = \ell\left(R_{\mu}(a) + \sum_{n = 1}^{\infty} (\mu - \lambda)^n R_{\mu}(a)^{n+1}\right)
    \end{equation}
    by \cref{lem:resolvent_set_open}. Since the series converges to \((\mu \unity - a)R_{\lambda}(a) - \unity\) uniformly and \(\ell\) is uniformly continuous, we have
    \begin{equation*}
        f_\ell(\lambda) = \ell(R_{\mu}(a)) + \sum_{n = 1}^\infty (-1)^n\ell\left(R_\mu(a)^{n+1}\right)(\lambda - \mu)^n,
    \end{equation*}
    which converges absolutely in \(B_{\norm{R_\mu(a)}^{-1}}\). That is, \(f_\ell\) is holomorphic in \(B_{\norm{R_\mu(a)}^{-1}}\), hence it can be extended to the connected component of \(\rho(a)\) containing \(\mu\).
\end{proof}

\begin{theorem}{Spectrum of an operator is compact}{spectrum_compact}
    Let \(\algebra{A}\) be a unital Banach algebra and \(a \in \algebra{A}\). Then \(\sigma(a)\) is non-empty and contained in \(\setc{\lambda \in \mathbb{C}}{\abs{\lambda} < \norm{a}}\).
\end{theorem}
\begin{proof}
    Suppose, by contradiction, \(\rho(a) = \mathbb{C}\), then by \cref{lem:holomorphic} the map \(f_\ell\) is entire for every \(\ell \in \algebra{A}^\dag\). Let \(\lambda \in \setc{\lambda \in \mathbb{C}}{\abs{\lambda} > \norm{a}}\), then
    \begin{equation*}
        R_{\lambda}(a) = \lambda^{-1}(\unity - \lambda^{-1}a)^{-1} = \lambda^{-1}\left(\unity + \sum_{n = 1}^{\infty} \lambda^{-n}u^n\right)
    \end{equation*}
    by \cref{thm:neumann_series_algebra}. We have
    \begin{equation*}
        \norm{R_{\lambda}(a)} \leq \abs{\lambda}^{-1}\left[1 + \sum_{n = 1}^\infty \left(\frac{\norm{u}}{\abs{\lambda}}\right)^n\right] = \abs{\lambda}^{-1} \left[1 + \frac{\frac{\norm{u}}{\abs{\lambda}}}{1 - \frac{\norm{u}}{\abs{\lambda}}}\right] = \frac{1}{\abs{\lambda} - \norm{u}}
    \end{equation*}
    which shows \(\norm{R_{\lambda}(a)} \to 0\) as we take \(\abs{\lambda} \to \infty\). Since \(\ell\) is continuous, we have \(\abs{f_\ell(\lambda)} \leq \norm{\ell}\cdot \norm{R_{\lambda}(a)}\), from which follows \(\abs{f_\ell(\lambda)} \to 0\) as \(\abs{\lambda} \to \infty\). Liouville's theorem ensures \(f_\ell\) is identically zero in the entire complex plane. This implies \(R_\lambda(a) \in \bigcap_{\ell \in \algebra{A}^\dag}\ker{\ell},\) hence \(R_{\lambda}(a) = 0\), which is not invertible. This contradiction shows \(\rho(a) \neq \mathbb{C}\), and we conclude \(\sigma(a)\) is non-empty. Moreover, we have shown \(\setc{\lambda \in \mathbb{C}}{\abs{\lambda} > \norm{a}} \subset \rho(a),\) hence \(\sigma(a) \subset \setc{\lambda \in \mathbb{C}}{\abs{\lambda} < \norm{a}}\).
\end{proof}
\begin{corollary}
    Let \(\algebra{A}\) be a unital Banach algebra and let \(a \in \algebra{A}\). Then \(\sigma(a)\) is compact.
\end{corollary}
\begin{proof}
    Since the complex plane has the Heine-Borel property, this result follows from the fact \(\norm{a}\) is a bound for the closed set \(\sigma(a)\).
\end{proof}

\subsection{Spectrum of projectors}
Abstracting the definition of projectors in Banach spaces and orthogonal projectors in Hilbert spaces we define projectors in general associative algebras.
\begin{definition}{Projectors and orthogonal projectors}{projector_algebra}
    Let \(\algebra{A}\) be an associative algebra. A \emph{projector} is an idempotent operator \(p \in \algebra{A}\), with \(p^2 = p\). If \(\algebra{A}\) is involutive, an \emph{orthogonal projector} is a self-adjoint projector.
\end{definition}

As expected from orthogonal projectors in Hilbert spaces, the norm of a orthogonal projector in C*-algebras is either zero or one.
\begin{proposition}{Norm of projectors}{norm_projector}
    Let \(\algebra{A}\) be a normed algebra. If \(p\in \algebra{A}\) is a projector, then either \(p = 0\) or \(\norm{p} \geq 1\). In addition, if \(\algebra{A}\) is a C*-algebra and \(p\) is an orthogonal projector, then \(\norm{p} \in \set{0,1}.\)
\end{proposition}
\begin{proof}
    In a normed algebra, we have \( \norm{p} = \norm{p^2} \leq \norm{p}^2\), which yields \(\norm{p} \in \set{0}\cup(1,\infty)\). It follows from the C* property and self-adjointness of the orthogonal projector that
    \begin{equation*}
        \norm{p} = \norm{p^2} = \norm{p^*p} = \norm{p}^2,
    \end{equation*}
    hence \(\norm{p} \in \set{0,1}\).
\end{proof}

\begin{proposition}{Necessary and sufficient condition for a projector}{projector_sufficient}
    Let \(\algebra{A}\) be a unital associative algebra. An operator \(p \in \algebra{A}\) is a projector if and only if \(\unity - p\) is a projector.
\end{proposition}
\begin{proof}
    If \(p\) is a projector, then
    \begin{equation*}
        (\unity - p)^2 = \unity - 2p + p^2  = \unity - p,
    \end{equation*}
    that is, \(\unity - p\) is a projector.

    Suppose \(\unity - p\) is a projector, then
    \begin{equation*}
        \unity - p=(\unity - p)^2 = \unity - 2p + p^2 \implies p^2 = p
    \end{equation*}
    that is, \(p\) is a projector.
\end{proof}
\begin{proposition}{Identity is the only invertible projector}{projector_invertible}
    Let \(\algebra{A}\) be a unital associative algebra. A projector \(p \in \algebra{A}\) is invertible if and only if \(p = \unity\).
\end{proposition}
\begin{proof}
    It is clear that \(\unity\) is a projector, since \(\unity^2 = \unity\). Suppose \(p \in \invertible{\algebra{A}}\) is a projector, then
    \begin{equation*}
        p^{-1}p^2 = p^{-1}p \implies p = \unity,
    \end{equation*}
    as desired.
\end{proof}

Let \(\algebra{A}\) be a unital Banach algebra, \(p \in \algebra{A}\) a projector, and \(\lambda \in \setc{\alpha \in \mathbb{C}}{\abs{\alpha} > 1}.\) Then \(\norm{\lambda^{-1}p} < 1\) and
\begin{equation*}
    \left(\unity - \lambda^{-1}p\right)^{-1} = \unity + \sum_{k = 1}^{\infty} \lambda^{-n}p^n  = \unity + \left(\sum_{k=1}^\infty \lambda^{-n}\right)p =\unity + \frac{1}{\lambda - 1}p
\end{equation*}
by \cref{thm:neumann_series_algebra}. \todo[Meromorphic map extension]
\begin{lemma}{Resolvent of a projector}{resolvent_projector}
    Let \(\algebra{A}\) be a unital associative algebra. If \(p\in\algebra{A}\) is a projector, then
    \begin{equation*}
        R_{\lambda}(p) = \frac{1}{\lambda}\unity + \frac{1}{\lambda(\lambda - 1)}p
    \end{equation*}
    is the resolvent of \(p\) at \(\lambda \in \mathbb{C} \setminus \set{0,1}.\)
\end{lemma}
\begin{proof}
    Let \(\lambda \in \mathbb{C}\setminus\set{0,1}\), then
    \begin{align*}
        \left[\frac1\lambda\unity + \frac1{\lambda(\lambda - 1)}p\right](\lambda \unity - p)
        &= \unity + \left(\frac{1}{\lambda-1} - \frac{1}{\lambda}\right)p - \frac{1}{\lambda(\lambda - 1)}p^2\\&= \unity + \left[\left(\frac{1}{\lambda-1} - \frac{1}{\lambda}\right) - \frac{1}{\lambda(\lambda - 1)}\right]p = \unity
    \end{align*}
    and
    \begin{align*}
        (\lambda \unity - p)\left[\frac1\lambda\unity + \frac1{\lambda(\lambda - 1)}p\right]
        &= \unity + \left(\frac{1}{\lambda-1} - \frac{1}{\lambda}\right)p - \frac{1}{\lambda(\lambda - 1)}p^2 \\&= \unity + \left[\left(\frac{1}{\lambda-1} - \frac{1}{\lambda}\right) - \frac{1}{\lambda(\lambda - 1)}\right]p = \unity,
    \end{align*}
    hence \(\lambda \in \rho(p)\) and the result follows.
\end{proof}

\begin{proposition}{Spectra of projectors}{spectra_projector}
    Let \(\algebra{A}\) be a unital associative algebra. The spectrum of a projector \(p \in \algebra{A}\) is contained in the set \(\set{0,1}\), with \(\sigma(p) = \set{0}\) if and only if \(p = 0\) and \(\sigma(p) = \set{1}\) if and only if \(p = \unity\).
\end{proposition}
\begin{proof}
    \cref{lem:resolvent_projector} guarantees \(\mathbb{C} \setminus \set{0,1}\subset \rho(p)\), hence \(\sigma(p) \subset \set{0,1}\) for any projector.

    In particular, \(0\) is not invertible, but \(\lambda \unity\) is invertible for all \(\lambda \in \mathbb{C} \setminus \set{0},\) thus showing that \(\sigma(0) = \set{0},\) and \(\unity\) is invertible with \((\lambda - 1)\unity\) invertible for all \(\lambda \in \mathbb{C}\setminus\set{1},\) that is, \(\sigma(\unity) = \set{1}\).

    Suppose \(\sigma(p) = \set{0},\) then \(\unity - p \in \invertible{\algebra{A}}\), hence \(p = 0\) by \cref{prop:projector_invertible}. Suppose \(\sigma(p) = \set{1}\), then \(-p \in \invertible{\algebra{A}}\), hence \(p = \unity\).
\end{proof}

\subsection{Spectral mapping theorem}
We consider a polynomial \(\varphi : \mathbb{C} \to \mathbb{C}\) of degree \(n \in \mathbb{N}\), then we may write
\begin{equation*}
    \varphi(z) = a_0 + \sum_{k = 1}^n a_k z^k
    \quad\text{and}\quad
    \varphi(z) = a_n \prod_{k = 1}^n (z - \alpha_k)
\end{equation*}
for all \(z \in \mathbb{C}\), where \(\ffamily{a_k}{k = 0}{n}\subset \mathbb{C} \) are the \emph{coefficients} of \(\varphi\) and \(\ffamily{\alpha_k}{k=1}{n} \subset \mathbb{C}\) are the \emph{roots} of \(\varphi\). In a unital associative algebra \(\algebra{A}\), we may define the polynomial
\begin{align*}
    \varphi_{\algebra{A}} : \algebra{A} &\to \algebra{A}\\
    u &\mapsto a_0 \unity + \sum_{k=1}^{n} a_k u^k
\end{align*}
by the coefficients of the complex polynomial \(\varphi\) and as a result we have also
\begin{equation*}
    \varphi_\algebra{A}(u) = a_n \prod_{k=1}^n (u - \alpha_k \unity)
\end{equation*}
for all \(u \in \algebra{A}\), since any operator commutes with its powers and with the identity \(\unity\). In this context, the constant polynomial \(\mathbb{C} \ni z \mapsto a_0 \in \mathbb{C}\) becomes the constant map \(\algebra{A} \ni u \mapsto a_0\unity \in \algebra{A}\).
\begin{proposition}{Operator commutes with its image under a polynomial}{operator_commutes_polynomial}
    Let \(\algebra{A}\) be a unital associative algebra and \(u \in \algebra{A}\). If \(w \in \algebra{A}\) satisfies \(w u = uw\), then
    \begin{equation*}
        \varphi_{\algebra{A}}(u)w = w\varphi_{\algebra{A}}(u)
    \end{equation*}
    for any polynomial \(\varphi\).
\end{proposition}
\begin{proof}
    By induction it is clear that \(u^kw = wu^k\) for any \(k \in \mathbb{N}\). Indeed, it holds for \(k = 1\) and supposing it holds for some \(m \in \mathbb{N}\), we have
    \begin{equation*}
        u^{m+1}w = u^{m}uw = u^mwu = wu^{m}u = wu^{m+1},
    \end{equation*}
    and it holds for \(m + 1\). The principle of finite induction ensures our claim holds for any \(k \in \mathbb{N}\).

    If \(\varphi\) is a constant polynomial, the result follows since \(w\) commutes with the identity and with the zero operator, so we may assume \(\varphi\) is a polynomial of degree \(n \in \mathbb{N}\). There exist coefficients \(a_k \in \mathbb{C}\) such that
    \begin{equation*}
        \varphi_{\algebra{A}}(u) = a_0 \unity + \sum_{k=1}^n a_k u^k,
    \end{equation*}
    hence
    \begin{equation*}
        w\varphi_{\algebra{A}}(u) = a_0 w + \sum_{k = 1}^n a_k wu^k = a_0 w + \left(\sum_{k = 1}^n a_{k}u^{k}\right)w = \left(a_0 \unity + \sum_{k=1}^n a_k u^k\right) w= \varphi_{\algebra{A}}(u) w,
    \end{equation*}
    as claimed.
\end{proof}

This construction allows us to relate the spectra of operators that are related by the image of such a polynomial. We'll write
\begin{equation*}
    \varphi(\sigma(u)) = \setc{\mu \in \mathbb{C}}{\exists \lambda \in \sigma(u) : \mu = \varphi(\lambda)}
\end{equation*}
and we'll establish the equality of this set with the spectrum of \(\varphi_{\algebra{A}}(u)\),
\begin{equation*}
    \sigma(\varphi_\algebra{A}(u)) = \setc{\mu \in \mathbb{C}}{\mu \unity - \varphi_\algebra{A}(u) \notin \invertible{\algebra{A}}},
\end{equation*}
result known as the spectral mapping theorem.
As a shorthand, we will simply denote the polynomial on the algebra by the same symbol as the complex polynomial, relying on the context to distinguish between them.
\begin{lemma}{Invertibility and polynomials}{polynomial_spectrum}
    Let \(\algebra{A}\) be a unital associative algebra and let \(u \in \algebra{A}\). If \(\lambda \in \sigma(u)\), then \((u - \lambda \unity)\psi(u) \notin \invertible{\algebra{A}}\) for any polynomial \(\psi\).
\end{lemma}
\begin{proof}
    If \(\psi\) is the trivial polynomial, the result follows trivially, so we may assume \(\psi\) has degree \(n \in \mathbb{N}\). Suppose, by contradiction, \(v = (u - \lambda \unity)\psi(u)\) is invertible for some \(\lambda \in \sigma(u)\), then there exists \(w \in \algebra{A}\) such that \(wv = vw = \unity\). Notice \(v\) commutes with \(u\) by \cref{prop:operator_commutes_polynomial}, then
    \begin{equation*}
        v (uw - wu) = uvw - vw u = u - u = 0 \implies w v( uw - wu) = 0 \implies uw = wu,
    \end{equation*}
    that is, \(w\) commutes with \(u\), hence it commutes with \(\psi(u)\) and \(u - \lambda \unity\). We conclude
    \begin{equation*}
        \unity = wv = w(u - \lambda \unity) \psi(u) = w \psi(u) (u - \lambda \unity)
    \end{equation*}
    and
    \begin{equation*}
        \unity = vw = (u - \lambda \unity) \psi(u) w = (u - \lambda \unity)w \psi(u),
    \end{equation*}
    that is, \((u - \lambda \unity) \in \invertible{\algebra{A}}\), hence \(\lambda \in \rho(u)\cap \sigma(u) = \emptyset\). This contradiction shows \((u - \lambda \unity)\psi(u)\) is not invertible.
\end{proof}
\begin{theorem}{Spectral mapping theorem}{spectral_mapping}
    Let \(\algebra{A}\) be a unital associative algebra and let \(u \in \algebra{A}\). Then
    \begin{equation*}
        \sigma(\varphi(u)) = \varphi(\sigma(u))
    \end{equation*}
    for any polynomial \(\varphi\).
\end{theorem}
\begin{proof}
    Consider a constant polynomial \(\varphi\) with constant term \(a_0 \in \mathbb{C}\), then \(\sigma(\varphi(u)) = \sigma(a_0\unity) = \set{a_0}\) and \(\varphi(\sigma(u)) = \set{a_0}\), hence \(\sigma(\varphi(u)) = \varphi(\sigma(u))\). We now suppose \(\varphi\) is a polynomial with degree \(n \in \mathbb{N}\).

    Let \(\mu \in \varphi(\sigma(u))\), then there exists \(\lambda \in \sigma(u)\) such that \(\varphi(\lambda) = \mu\). We consider the polynomial \(\varphi(z) - \mu\), which has \(\lambda\) as a root, hence there exists a polynomial \(\psi\) such that \(\varphi(z) - \mu = (z - \lambda)\psi(z)\). We then have \(\varphi(u) - \mu \unity = (u - \lambda \unity)\psi(u)\), that is, \(\varphi(u) - \mu \unity \notin \invertible{\algebra{A}}\) by \cref{lem:polynomial_spectrum}. This shows \(\mu \in \sigma(\varphi(u))\), hence \(\varphi(\sigma(u)) \subset \sigma(\varphi(u))\).

    Let \(\mu \in \sigma(\varphi(u))\), and consider the polynomial \(\varphi(u) - \mu\unity = a_n \prod_{k=1}^n (u - \alpha_k\unity)\), where \(\alpha_k\) are the roots of the polynomial \(\varphi(z) - \mu\) and \(a_n \neq 0\) the \(n\)th term coefficient of \(\varphi\). The polynomial is not invertible by hypothesis, hence there exists \(m \in \set{1,2,\dots, n}\) such that \(\alpha_m \in \sigma(u)\), otherwise \(u - \alpha_k \unity \in \invertible{\algebra{A}}\) for all \(k \in \set{1,2, \dots, n}\) and \(\varphi(u) - \mu\) would be invertible. We have \(\varphi(\alpha_m) - \mu = 0,\) and we conclude \(\mu \in \varphi(\sigma(u))\), that is, \(\sigma(\varphi(u)) \subset \varphi(\sigma(u))\).
\end{proof}

\subsection{Spectral radius}
In \cref{thm:spectrum_compact}, we've shown that in a unital Banach algebra, the spectrum of an operator is contained in an open disk centered at zero of radius equal to its norm.
\begin{definition}{Spectral radius}{spectral_radius}
    Let \(\algebra{A}\) be a unital algebra. For an operator \(a \in \algebra{A}\), we define
    \begin{equation*}
        r(a) = \sup_{\lambda \in \sigma(a)}{\abs{\lambda}}
    \end{equation*}
    as is its \emph{spectral radius}.
\end{definition}
In a Banach algebra \(\algebra{A}\), we have \(r(a) \leq \norm{a}\) for all \(a \in \algebra{A}\), that is, the spectrum of an operator is contained in the closed disk centered at zero of radius equal to its spectral radius.

\begin{theorem}{Spectral radius theorem}{spectral_radius_theorem}
    If \(\algebra{A}\) is a unital Banach algebra, then
    \begin{equation*}
        r(a) = \inf_{n\in \mathbb{N}}{\norm{a^n}^{\frac{1}{n}}}= \lim_{n\to\infty}{\norm{a^n}^{\frac1n}}
    \end{equation*}
    for all \(a \in \algebra{A}\).
\end{theorem}
\begin{proof}
    From \cref{thm:spectral_mapping}, we know that if \(\lambda \in \sigma(a)\), then \(\lambda^n \in \sigma(a^n)\) for all \(n \in \mathbb{N}\). The boundedness of spectra yields \(\abs{\lambda^n} \leq \norm{a^n}\) for all \(n \in \mathbb{N}\), hence
    \begin{equation*}
        r(a) = \sup_{\lambda \in \sigma(a)}{\abs{\lambda}}\leq\inf_{n\in \mathbb{N}}{\norm{a^n}^{\frac1n}} \leq \sup_{n \in \mathbb{N}}{\inf_{m \in \mathbb{N}}{\setc*{\norm{a^m}^{\frac1m}}{m \geq n}}} = \liminf_{n\to\infty}{\norm{a^n}^{\frac1n}}.
    \end{equation*}

    Let \(D = \setc{z \in \mathbb{C}}{\abs{z} \leq r(a)},\) then \(\sigma(a) \subset D\) and \(\mathbb{C} \setminus D\) is a connected component of \(\rho(a)\). As shown in \cref{lem:holomorphic}, for every \(\ell \in \algebra{A}^\dag\), the map \(f_\ell : \mathbb{C}\setminus D \to \mathbb{C}\) defined by \(\lambda \mapsto \ell(R_{\lambda}(a))\) is holomorphic in \(\mathbb{C} \setminus D\), hence there exist \(\alpha_k \in \mathbb{C}\) with \(k \in \mathbb{Z}\) such that
    \begin{equation*}
        f_\ell(\lambda) = \sum_{k \in \mathbb{Z}} \alpha_k \lambda^{-k}
    \end{equation*}
    is the Laurent series for \(f_\ell\) in \(\mathbb{C} \setminus D\). For all \(\lambda \in \setc{z \in \mathbb{C}}{\abs{z} > \norm{a}} \subset \mathbb{C} \setminus D\), we have \(\norm{\lambda^{-1}a} < 1\), then the \nameref{thm:neumann_series_algebra} yields
    \begin{equation*}
        f_\ell(\lambda) = \ell\left((\lambda \unity - a)^{-1}\right)= \lambda^{-1} \ell\left((\unity - \lambda^{-1}a)^{-1}\right) = \lambda^{-1} \ell\left[\unity + \sum_{k = 1}^\infty (\lambda^{-1}a)^k\right] = \ell(\unity)\lambda^{-1} + \sum_{k = 1}^\infty \ell(a^k)\lambda^{-1 - k},
    \end{equation*}
    since \(\ell\) is uniformly continuous. We may conclude \(\alpha_1 = \ell(\unity),\) \(\alpha_{n+1} = \ell(a^n)\) and \(\alpha_{-n} = 0\) for all \(n \in \mathbb{N}\), that is,
    \begin{equation*}
        f_\ell(\lambda) = \ell(\unity)\lambda^{-1} + \sum_{n = 1}^\infty \ell(a^n) \lambda^{-1 - n}
    \end{equation*}
    for all \(\lambda \in \mathbb{C} \setminus D\). In particular, the series converges, hence
    \begin{equation*}
        \lim_{n\to \infty}{\abs*{\ell(a^n) \lambda^{-1-n}}} = 0,
    \end{equation*}
    and we may conclude the sequence \(\family{\ell(a^n)\lambda^{-1-n}}{n \in \mathbb{N}}\) converges, thus bounded. Let
    \begin{equation*}
        \mathcal{T} = \setc{\lambda^{-1-n} \eval(a^n)}{n \in \mathbb{N}, \lambda \in \mathbb{C}\setminus D} \subset (\algebra{A}^\dag)^\dag,
    \end{equation*}
    then we have thus shown the set \(\setc{\abs{T\ell}}{T \in \mathcal{T}}\) is bounded for every \(\ell \in \algebra{A}^\dag\). The \nameref{thm:Banach_Steinhaus} ensures that \(\setc{\norm{T}}{T \in \mathcal{T}}\) is bounded by some \(M \geq 0\). Since the evaluation map is an isometry, we have \(\norm{\lambda^{-1-n}a^n} \leq M\) for all \(n \in \mathbb{N}\) and all \(\lambda \in \mathbb{C} \setminus D\). This yields
    \begin{equation*}
        \norm{a^n}^{\frac1n} \leq M^{\frac1n} \abs{\lambda}^{\frac1n} \abs{\lambda}
    \end{equation*}
    for all \(n \in \mathbb{N}\) and \(\lambda \in \mathbb{C} \setminus D\). It is a standard result in real analysis that \(x^{\frac1n} \to 1\) for any \(x > 0\), therefore
    \begin{equation*}
        \norm{a^n}^{\frac1n} \leq \inf_{\lambda \in \mathbb{C} \setminus D}{\abs{\lambda}} = r(a)
    \end{equation*}
    for all \(n \in \mathbb{N}\). In particular, this yields
    \begin{equation*}
        \limsup_{n\to \infty}{\norm{a^n}^{\frac1n}} \leq r(a) \leq \inf_{n \in \mathbb{N}}{\norm{a^n}^{\frac1n}} \leq \liminf_{n\to\infty}{\norm{a^n}^{\frac1n}},
    \end{equation*}
    hence
    \begin{equation*}
        \lim_{n\to\infty}{\norm{a^n}^{\frac1n}} = r(a),
    \end{equation*}
    as desired.
\end{proof}

\begin{definition}{Quasi-nilpotent operator}{quasi_nilpotent}
    Let \(X\) be a Banach space. A bounded operator \(A \in \bounded(X)\) that satisfies
    \begin{equation*}
        \lim_{n\to \infty} \norm{A^n}^{\frac1n} = 0
    \end{equation*}
    is said to be \emph{quasi-nilpotent}.
\end{definition}
\begin{remark}
    By the spectral radius theorem, we conclude the spectrum of a quasi-nilpotent operator is always \(\set{0}\). Indeed, if \(A\) is quasi-nilpotent, then \(r(A) = 0\), which shows us \(\sigma(A) = \set{0}\).
\end{remark}

\begin{example}{Volterra operator in \(\mathcal{C}([0, 1], \mathbb{R})\)}{volterra}
    Let \(X = \mathcal{C}([0,1], \mathbb{R})\) be equipped with the supremum norm and let \(W : X \to X\) be the \emph{Volterra operator} defined by
    \begin{equation*}
        Wf(x) = \int_0^x \dli{\xi} f(\xi)
    \end{equation*}
    for all \(f \in X\) and \(x \in [0,1]\). Then \(W\) is quasi-nilpotent and \(\norm{W} = 1\).
\end{example}
\begin{proof}
    Let \(f \in X\), then for all \(\xi \in [0,1]\) we have \(\abs{f(\xi)} \leq \norm{f}_\infty\). This yields
    \begin{equation*}
        \abs{Wf(x)} \leq \int_0^x \dli{\xi} \abs{f(\xi)} \leq x \norm{f}_\infty
    \end{equation*}
    for all \(x \in [0,1]\), hence \(\norm{Wf}_\infty \leq \norm{f}_\infty\), which shows \(W\) is bounded with \(\norm{W} \leq 1\). By the same argument, we also have \(\abs{W1(x)} \leq x\) for all \(x \in [0,1]\), where \(1 \in X\) is the constant function \(1(x) = 1\). Then \(\norm{W1}_\infty = 1\) and \(\norm{1}_\infty = 1\), from which follows \(\norm{W} \geq 1\), and we conclude \(\norm{W} = 1\).

    Consider the set
    \begin{equation*}
        S = \setc*{n \in \mathbb{N}}{\forall x \in [0,1], \forall f \in X : \abs*{(W^nf)(x)} \leq \frac{x^n}{n!} \norm{f}_\infty},
    \end{equation*}
    then we have just shown \(1 \in S\) and, in particular, that \(S\) is non-empty. Let \(k \in S\), then for all \(f \in X\) and \(x \in [0,1]\) we have
    \begin{equation*}
        \abs*{(W^{k+1}f)(x)} \leq \int_0^x \dli{\xi} \abs*{(W^kf)(\xi)} \leq \int_0^x \dli{\xi} \frac{\xi^k}{k!}\norm{f}_\infty = \frac{x^{k+1}}{(k+1)!}\norm{f}_\infty,
    \end{equation*}
    that is, \(k + 1 \in S\). By the principle of finite induction, we have shown that \(S = \mathbb{N}\). As a result, we have \(\norm{W^n f}_\infty \leq \frac{\norm{f}_\infty}{n!}\), which yields \(\norm{W^n} \leq \frac{1}{n!}\) for all \(n \in \mathbb{N}\).

    For every \(n \in \mathbb{N}\) we have
    \begin{equation*}
        n! = \prod_{k = 1}^n k > \prod_{k = \ceil*{\frac{n}{2}}}^n k > \prod_{k = \ceil*{\frac{n}{2}}}^{n} \ceil*{\frac{n}{2}} = \ceil*{\frac{n}{2}}^{\ceil*{\frac{n}{2}}} > \left(\frac{n}{2}\right)^{\frac{n}{2}},
    \end{equation*}
    then
    \begin{equation*}
        0 < \frac{1}{n!} <  \left(\frac{2}{n}\right)^{\frac{n}{2}} \implies 0 < \left(\frac{1}{n!}\right)^{\frac1n} < \left(\frac{2}{n}\right)^{\frac12}.
    \end{equation*}
    Since the left and right hand sides converge to \(0\) as \(n \to \infty\), we have shown
    \begin{equation*}
        r(W) = \lim_{n\to\infty}{\norm{W^n}^{\frac1n}} = \lim_{n\to \infty}{\left(\frac{1}{n!}\right)^{\frac1n}} = 0,
    \end{equation*}
    that is, \(W\) is quasi-nilpotent.
\end{proof}

\begin{theorem}{Spectral radius for self-adjoint or normal operators in C*-algebras}{spectral_radius_cstar}
    Let \(\algebra{A}\) be a unital C*-algebra. If \(a \in \setc{u \in \algebra{A}}{u^* = u \lor u^*u = uu^*}\), that is, if \(a\) is either a self-adjoint or a normal operator, then \(\norm{a} = r(a)\).
\end{theorem}
\begin{proof}
    Let \(a \in \algebra{A}\) be a self-adjoint operator. Let us show by induction that for all \(n \in \mathbb{N}\) we have \(\norm{a^{(2^n)}} = \norm{a}^{(2^n)}\). Indeed, for \(n = 1\) we have \(\norm{a}^2 = \norm{a^*a} = \norm{a^2}\). Suppose it holds for some \(k \in \mathbb{N}\), then
    \begin{equation*}
        \norm{a^{(2^{k+1})}} = \norm{[a^{(2^k)}]^*a^{(2^k)}}= \norm{a^{(2^k)}}^2 = [\norm{a}^{(2^k)}]^2 = \norm{a}^{(2^{k+1})},
    \end{equation*}
    therefore it holds for \(k + 1 \in \mathbb{N}\). By the principle of finite induction we conclude it holds for all \(n \in \mathbb{N}\). We consider the sequence \(\family{\norm{a^n}^{\frac1n}}{n \in \mathbb{N}}\) which converges to \(r(a)\). By \cref{prop:convergence_subsequence} the subsequence \(\family{\norm{a^{(2^n)}}^{\frac1{2^n}}}{n \in \mathbb{N}}\) converges to \(r(a)\), which yields
    \begin{equation*}
        r(a) = \lim_{n \to \infty}{\norm{a^{(2^n)}}^{\frac1{2^n}}} = \lim_{n\to\infty}{\norm{a}} = \norm{a}.
    \end{equation*}

    Let \(b \in \algebra{A}\) be a normal operator, then \((b^*b)^{(2^{n})} = [b^{(2^n)}]^*b^{(2^n)}\) for all \(n \in \mathbb{N}\). Indeed, we have \((b^*b)^2 = b^*bb^*b = (b^*)^2 b^2\) and supposing true for some \(n \in \mathbb{N}\) we have
    \begin{equation*}
        [b^{(2^{n+1})}]^* b^{(2^{n+1})} = [b^{(2^n)}]^* [b^{(2^n)}]^* b^{(2^n)} b^{(2^n)} = [b^{(2^n)}]^*  b^{(2^n)} [b^{(2^n)}]^*b^{(2^n)} = [(b^*b)^{(2^n)}]^2 = (b^*b)^{(2^{n+1})},
    \end{equation*}
    that is, it also holds for \(n + 1 \in \mathbb{N}\). The claim is true for all \(n \in \mathbb{N}\) by the principle of finite induction. Since \(b^*b\) is self-adjoint, we have by a previous result and the C*-property that
    \begin{equation*}
        \norm{(b^*b)^{(2^n)}}= \norm{b^*b}^{(2^n)} = \norm{b}^{(2^{n+1})}
    \end{equation*}
    for all \(n \in \mathbb{N}.\) Repeating the subsequence argument yields \(r(b) = \norm{b}\).
\end{proof}

We may use \cref{thm:spectral_mapping,thm:spectral_radius_cstar} to show a result reminiscent of Cayley-Hamilton theorem for linear operators in finite dimensional linear spaces.
\begin{proposition}{Operator with finite and real spectrum}{cayley_hamilton}
    Let \(\algebra{A}\) be a unital C*-algebra and let \(a \in \algebra{A}\) be a self-adjoint operator. Suppose \(\sigma(a) \subset \mathbb{R}\) is finite, and let \(\varphi : \mathbb{C} \to \mathbb{C}\) be a polynomial with zeros given by \(\sigma(a)\), then \(\varphi(a) = 0\).
\end{proposition}
\begin{proof}
    We may assume without loss of generality that
    \begin{equation*}
        \varphi(z) = \prod_{\lambda \in \sigma(a)} (z - \lambda),
    \end{equation*}
    since \(\alpha \varphi\) satisfies \(\alpha \varphi(a) = 0\) if and only if \(\varphi(a) = 0\), for any \(\alpha \in \mathbb{C}\setminus \set{0}\). Since the spectrum is real, \(\varphi\) is a polynomial with real coefficients, hence \(\varphi(a)\) is self-adjoint as a real linear combination of self-adjoint operators. By \cref{thm:spectral_mapping} we have \(\sigma(\varphi(a)) = \set{0}\) and we conclude \(\norm{\varphi(a)} = 0\) from \cref{thm:spectral_radius_cstar}.
\end{proof}
\begin{corollary}
    Let \(\algebra{A}\) be a unital C*-algebra. If \(a \in \algebra{A}\) is self-adjoint and \(\sigma(a) \subset \set{-1,1}\), then \(a^2 = \unity\).
\end{corollary}
\begin{proof}
    We consider the polynomial \(z^2 - 1\) whose zeros is a subset of \(\sigma(a)\), then by \cref{prop:cayley_hamilton} yields \(a^2 - 1 = 0\).
\end{proof}

Another consequence of the spectral radius theorem is that it algebraically defines the operator norm in a C*-algebra.
\begin{proposition}{Spectral radius defines the norm in a C*-algebra}{norm_cstar_spectral_radius}
    Let \(\algebra{A}\) be a unital C*-algebra. Then
    \begin{equation*}
        \norm{a} = \sqrt{r(a^*a)}
    \end{equation*}
    for all \(a \in \algebra{A}\).
\end{proposition}
\begin{proof}
    Let \(a \in \algebra{A}\), then \(a^*a\) is self-adjoint. By \cref{thm:spectral_radius_cstar}, we have \(\norm{a^*a} = r(a^*a),\) and the result follows from the C* property.
\end{proof}
\begin{corollary}
    Let \(\algebra{A}\) be a unital *-algebra. If \(\norm{\noarg}_1 : \algebra{A} \to [0,\infty]\) and \(\norm{\noarg}_2 : \algebra{A} \to [0, \infty]\) are norms in \(\algebra{A}\) such that \((\algebra{A}, \norm{\noarg}_1)\) and \((\algebra{A},\norm{\noarg}_2)\) are C*-algebras, then \(\norm{\noarg}_1 = \norm{\noarg}_2\).
\end{corollary}
\begin{proof}
    Indeed, for any operator \(a \in \algebra{A}\) we have \(\norm{a}_1 = \sqrt{r(a^*a)} = \norm{a}_2\).
\end{proof}

\begin{proposition}{Spectra of unitary operators lie in the unit circle}{spectra_unitary}
    Let \(\algebra{A}\) be a unital C*-algebra. If \(a \in \algebra{A}\) is a unitary operator, then \(\sigma(a) \subset S^1 = \setc{z \in \mathbb{C}}{\abs{z} = 1}\).
\end{proposition}
\begin{proof}
    Notice \(a\) is normal, since \(aa^* = \unity = a^*a\), hence \(\norm{a} = 1\) by \cref{thm:spectral_radius_cstar}. Since \(r(a) = 1\), we have \(\sigma(a) \subset D = \setc{z \in \mathbb{C}}{\abs{z} \leq 1}\). This also holds for \(a^{-1} = a^*\), that is, \(\sigma(a^*) = \sigma(a^{-1}) \subset D\), with \(\conj{\sigma(a^*)}= \sigma(a) = \sigma(a^{-1})^{-1}\). This shows every \(\lambda \in \sigma(a)\) satisfies \(\abs{\lambda} \leq 1\), \(\lambda^{-1} \in \sigma(a)\), and \(\conj{\lambda} \in \sigma(a)\) and therefore \(\lambda \in S^1\). Indeed, were it not the case, we would have \(\lambda^{-1} \in \mathbb{C} \setminus D\), which contradicts \(\sigma(a^{-1}) \subset D\).
\end{proof}

\begin{theorem}{Spectra of self-adjoint operators is a compact subset of \(\mathbb{R}\)}{spectra_self_adjoint}
    Let \(\algebra{A}\) be a unital C*-algebra. If \(a \in \algebra{A}\) is self-adjoint, then \(\sigma(a)\) is a compact subset of \([-\norm{a}, \norm{a}]\).
\end{theorem}
\begin{proof}
    We may assume without loss of generality \(a\neq0\), since \(\sigma(0) = \set{0}\) is a compact subset of \(\set{0}\).
    Let \(\mu \in (0, \norm{a}^{-1})\), then \(\norm{\pm i\mu a} = \mu \norm{a} < 1\) and it follows from \cref{thm:neumann_series_algebra} that \(\unity \pm i \mu a \in \invertible{\algebra{A}}\).

    We consider \(R = \setc{z \in \mathbb{C}}{\abs{z} \leq \norm{a} \land \Im(z) \neq 0}\) and we aim to show \(R \subset \rho(a)\). Let \(\lambda \in R\), then \(1 \pm i \mu \lambda \neq 0\) since \(\mu < \norm{a}^{-1} \leq \abs{\lambda}^{-1}\). Then
    \begin{align*}
        \lambda \unity - a &= \frac{(1 + i \mu \lambda) - (1 - i \mu \lambda)}{2i \mu}\unity - \frac{i\mu\left[(1 + i\mu \lambda) + (1 - i \mu \lambda)\right]}{2i \mu} a\\
                           &= \frac{1}{2i\mu}\left[(1 + i\mu \lambda)(\unity - i\mu a) - (1 - i\mu \lambda)(\unity + i\mu a)\right]\\
                           &= \frac{1 - i\mu \lambda}{2i\mu} \left[\frac{1 + i\mu \lambda}{1 - i\mu \lambda}(\unity - i\mu a) - (\unity + i\mu a)\right]\\
                           &= \frac{1 - i\mu \lambda}{2i \mu} \left[\frac{1 + i\mu \lambda}{1 - i\mu \lambda}\unity - (\unity + i\mu a)(\unity - i\mu a)^{-1}\right](\unity - i\mu a).
    \end{align*}
    Let \(\kappa = \frac{1 + i\mu \lambda}{1 - i\mu \lambda}\), then
    \begin{equation*}
        \abs{\kappa}^2 = \frac{(1 + i\mu \lambda)(1 -i \mu \conj{\lambda})}{(1 - i\mu \lambda)(1 + i \mu \conj{\lambda})} = \frac{1 + \mu^2 \abs{\lambda}^2 - 2\mu \Im{\lambda}}{1 + \mu^2 \abs{\lambda}^2 + 2\mu \Im{\lambda}} = 1 - \frac{4\mu\Im{\lambda}}{1 + \mu^2 \abs{\lambda}^2 + 2 \mu \Im{\lambda}} \neq 1,
    \end{equation*}
    since \(\Im \lambda \neq 0\). Let \(u = (\unity + i\mu a)(\unity - i\mu a)^{-1}\), then \(u \in \invertible{\algebra{A}}\) and we have
    \begin{equation*}
        u^* = \left[(\unity - i \mu a)^{-1}\right]^*(\unity - i\mu a) = (\unity + i \mu a)^{-1} (\unity - i \mu a) = (\unity - i \mu a)(\unity + i \mu a)^{-1} = u^{-1},
    \end{equation*}
    where we have used \cref{prop:resolvent_commute}. Since \(\kappa \notin S^1\), we have \(\kappa \in \rho(u)\) by \cref{prop:spectra_unitary}, hence
    \begin{equation*}
        \lambda \unity - a = \frac{1 - i \mu \lambda}{2i \mu} R_{\kappa}(u)(1 - i \mu a) \in \invertible{\algebra{A}},
    \end{equation*}
    that is, \(\lambda \in \rho(a)\). \cref{thm:spectrum_compact} states \(\sigma(a)\) is a compact subset of \(\setc{z \in \mathbb{C}}{\abs{z} \leq \norm{a}} = R \cup [-\norm{a}, \norm{a}]\). As \(\sigma(a) \cap R = \emptyset\), we have \(\sigma(a) \subset [-\norm{a}, \norm{a}]\).
\end{proof}
